Generality in food webs scales with species richness, not latitude

### 2015-02-02 manuscript/latitudepaper.tex
half-way through methods

%Food webs are a thing. (Mention gen, vul, LS.) %Food webs have strong scaling structure.
Food webs --networks of feeding links between species-- have been used for over a century to summarize the structure of 
ecological communities \citep{Williams2000,earlierwork} and to understand how that structure relates to environmental variables 
such as land use \citep{Townsend1998,Townsend2004c,Digel2014}, primary productivity \citep{}, and climate\citep{}. The latter
variables in turn have strong gradients over latitude, with productivity and temperature both being higher in the tropics while 
climate is more variable at high latitudes \citep{}, 
suggesting that latitude may indirectly affect food-web structure.
Indeed, these latitudinal environmental gradients have been put forward as potential driverd for the latitudinal gradient in species 
richness, one of the most general and robust patterns in ecology \citep{Schemske2009a,,}.

This effect of latitude on species richness, however, presents a problem for the analysis of other aspects of food-web structure.
Although it was initially thought that food-web structure was small-world and scale invariant \citep{}, further 
analyses with higher-quality data have revealed that most food-web structural properties scale strongly with species 
richness, connectance (the proportion of possible links that are realized), or both \citep{Riede2010} 
(see \citet{Dunne2006} for a review of the history of scaling laws in food web theory). This means that any relationship
between latitude and food-web structure must be considered in light of the latitudinal gradient in species richness.
This is especially true for food-web properties such as measures of specialization
that both scale with species richness \citep{Riede2010} and are related to candidate
explanations for the latitudinal diversity gradient \citep{Vazquez2004}.

[[Smoother transition needed here]]
% Explain niche space hypothesis, Schemske, Vazquez. Then make clear that niche space suggests opposite gradient to ~S^B.
%Then give hypotheses.
A latitudinal gradient in specialization lies behind the ``latitude-niche breadth hypothesis'' \citep{Vazquez2004}.
This hypothesis predicts that decreased seasonality in the tropics should lead to more stable populations, which in
turn should evolve smaller niches \citep{Vazquez2004}. These narrow niches should therefore allow more species to 
coexist in the tropics than can at higher latitudes. Alternatively, higher productivity of the tropics \citep{Brown2004}
may result in a broader niche space \citep{Davies2007} which could also allow greater diversification even if niche 
sizes are globally similar. Although the assumptions of the latitude-niche breadth hypothesis are only equivocably 
supported \citep{Vazquez2004}, it remains a compelling potential mechanism for the latitudinal gradient in species 
richness \citep{Lappalainen2006,Krasnov2008,Slove2010}.

Importantly, the expectation of the latitude-niche breadth hypothesis that higher species richness is associated with 
reduced competition is contary to the expected increase in the mean number of interactions per species (one potential 
measure of specialization) as species richness increases \citep{Dunne2006,Riede2010}. If, therefore, there is a
distinct effect of latitude on specialization then this effect should be visible in changing scaling relationships.
Here, we tested for changes in the scaling of mean links per species, mean generality, and mean vulnerability with species
richness over latitude. We focused on these three food-web properties --only some of the wide array of metrics that have been
described (see e.g. \citep{})-- because together they provide a measure a species' biotic specialization, which underlies the 
``latitude-niche breadth hypothesis''.

% In addition to species richness, there are general and robust latitudinal gradients in metabolic rates \citep{Stegen2012} and 
% rates and intensities of species interactions \citep{Marquis2005,Schemske2009}. These gradients, together with the strength and 
% generality of associations between biological processes and environmental variables, suggest that variation in network structure 
% may also be explained by ecological gradients \citep{Baiser2012}. Importantly, some of these gradients may act directly 
% on food-web properties such that they change over latitude beyond the expected changes due to changing species richness. For 
% example, if niche sizes vary with latitude then mean numbers of predators, prey, and links per species may also vary with 
% latitude. Some of this variation may be explained by variation in species richness, but as niche sizes are believed to be 
% related directly to latitude it is reasonable to expect that there may also be latitudinal effects unrelated to those of 
% changing species richness.

@@

We compiled a list of 263 empirical food webs from
multiple sources (see Supplemental Information for web origins and selection
criteria). We recorded study site latitude from the original source where
possible or, where study sites were described but exact coordinates were not
given, obtained estimated coordinates using Google Earth \citep{GoogleEarth}.
Webs were also classified according to ecosystem type (terrestrial, lake,
stream, estuary, or marine), year of publication (to control for potential change
in standards of data collection over time \citep{Dunne2006}), and whether or not
humans were included.

@@

We then converted each web to a trophic-species web by aggregating species with identical predator and prey sets.
Using trophic species webs helps to reduce variation in resolution across different studies and across taxa within
a study \citep{Martinez1991,Vermaat2009,Dunne2004,Dunne2013}. The number of species ($S$) and links ($L$) in each 
trophic species
web were counted and used to calculate mean links per species ($Z$), generality ($G$), and vulnerability ($V$).

@@

We first examined simple linear relationships between latitude and $S$, $Z$, $G$, and $V$. We fit models of the form

@@

where $S_{i}$ is the species richness of web $i$, $\Lambda_{i}$ its absolute
latitude (degrees north or south  regardless of direction), $E_{i}$ is the
ecosystem type of network $i$ and $\epsilon_{i}$ a residual error term. If the
interaction between latitude and ecosystem type was not significant, this term
was removed and the model re-fit. These simple linear models, fit using the
function lm in R \citep{R} package stats, were used to determine whether there
were latitudinal gradients in food-web structure in this data set.

@@

We were then able to examine the form of the scaling relationship between each property ($Z$, $G$, and $V$) and $S$.
The scaling relationship between $Z$ and $S$ has been shown to be a power law \citep{Riede2010} of the form

\begin{equation}
\label{Power}
Z_{i}=\alpha S_{i}^{\beta} + \epsilon_{i} 
\end{equation}

@@

\begin{equation}
\label{Loglog}
log(Z_{i}) = log(\alpha) + \beta log(S_{i}) + \epsilon_{i}  
\end{equation}

Although scaling relationships between $S$, $G$, and $V$ have not been explicitly examined (but see scaling 
relationships for the standard devaitions of each property in \citet{Riede2010}), we expect that they will follow
power laws similar to that of the relationship between $S$ and $Z$. This is because the links taken into account in
calculating $G$ and $V$ are subsets of the total links included when calculating $Z$. As with $Z$, we explicitly 
compared the error distributions of models for $G$ and $V$ using both the power-law and logarithmic formulations.

%%% TO SI

% Note that the two forms of the power-law relationship imply different error structures: normally-distributed, additive error in 
% \ref{Power} and lognormal, multiplicative error in \ref{Loglog}. As there is no intuitive reason to believe that our
% data have one error distribution over another, we follow the recommendations in \citet{Xiao2011} and compared the two
% model formulations explicitly. The model with the error distribution most resembling that observed in the empirical
% data was then used to test for potential effects of latitude.

>>>>

%Food webs are a thing. (Mention gen, vul, LS.) %Food webs have strong scaling structure.
Food webs --networks of feeding links between species-- have been used for over a century to summarize the structure of 
ecological communities \citep{Williams2000,earlierwork} and to understand how that structure relates to environmental variables 
such as land use \citep{Townsend1998,Townsend2004c,Digel2014}, primary productivity \citep{}, and climate\citep{}. The latter
variables in turn have strong gradients over latitude, with productivity and temperature both being higher in the tropics while 
climate is more variable at high latitudes \citep{}. These variables affect both the resources available and species' 
metabolisms~\citep{Stegen2012}, and have been proposed as influences on the strength of 
interspecific interactions~\citep{Marquis2005,Schemske2009}. 
By shaping interactions between species, latitudinal gradients may also shape food-web structure.
Indeed, these latitudinal environmental gradients have been put forward as potential driverd for the 
latitudinal gradient in species richness, one of the most general and robust patterns in ecology \citep{Schemske2009a,,}.

% These gradients, together with the strength and generality of associations between biological processes and environmental variables, suggest that variation in network structure 
% may also be explained by ecological gradients \citep{Baiser2012}.

One proposed link between species richness and latitude is the ``latitude-niche breadth hypothesis''~\citep{,Vazquez2004}. 
This hypothesis predicts that decreased seasonality in the tropics should lead to more stable populations, which in
turn should evolve smaller niches \citep{Vazquez2004}. These narrow niches should therefore allow more species to 
coexist in the tropics than can at higher latitudes. Alternatively, higher productivity of the tropics \citep{Brown2004}
may result in a broader niche space \citep{Davies2007} which could also allow greater diversification even if niche 
sizes are globally similar. Although the assumptions of the latitude-niche breadth hypothesis are only equivocably 
supported \citep{Vazquez2004}, it remains a compelling potential mechanism for the latitudinal gradient in species 
richness \citep{Lappalainen2006,Krasnov2008,Slove2010}.

If decreased seasonality or higher productivity is the cause of higher species-richness in the tropics, then 
there should be other effects of latitude on food-web structure. Attempts to unravel the effects, however,
are complicated by known effects of species richness on many other network properties~\citep{Riede2010}. In particular,
narrower niches imply fewer links per species (i.e., greater specialization in the tropics~\citep{}). However,
average numbers of links per species tend to increase in larger food webs~\citep{Dunne2006,Riede2010}. This means
that a latitudinal effect on specialization may be obscured by a latitudinal gradient in species richness. If this is
the case, it may still be possible to uncover effects of latitude on specialization by examining the shape of the scaling 
relationship between specialization and species richness over changing latitude. Here, we use three measures of specialization:
mean links per species, mean generality (number of prey), and mean vulnerability (number of predators). By testing whether
latitude affects the scaling of each property with species richness, we test effects of latitude on specialization implied by
the ``latitude-niche breadth hypothesis''.

@@

We compiled a list of 263 empirical food webs from
multiple sources (see Supplemental Information for web origins and selection
criteria). We recorded study site latitude from the original source where
possible or, where study sites were described but exact coordinates were not
given, obtained estimated coordinates using Google Earth \citep{GoogleEarth}.

@@

As the food webs in this dataset are derived from a variety of sources and were compiled over many decades, it
is likely that they vary in their resolution and in the amount of sampling effort invested in their assembly.
Many analyses of food-web structure reduce this variation by aggregating species with identical predator and prey
sets to form ``trophic species'' webs (e.g. ~\citep{Martinez1991,Vermaat2009,Dunne2004,Dunne2013}). As this study
is concerned directly with the number of species at a particular latitude, however, we did not wish to ignore 
species with redundant sets of interactions. We therefore analysed both original and trophic-species versions
of the dataset. As the version of the dataset used did not qualitatively change the results, we present only
the results for the original webs in the main text. In each case, we counted the number of species ($S$) and 
links in each web and used these to calculate the mean number of links per species ($Z$), mean generality 
($G$), and mean vulnerability ($V$) of each web.

@@

[[re-run with dredge]]
To determine whether there were latitudinal gradients in food-web structure,
we first examined simple linear relationships between latitude and $S$, $Z$, $G$, and $V$. 
We fit models of the form

@@

where $S_{i}$ is the species richness of web $i$, $L_{i}$ its absolute
latitude (degrees north or south  regardless of direction), $E_{i}$ is the
ecosystem type of network $i$ (comprising terms for stream, marine, lake, and terrestrial networks with estuarine
networks being the intercept) and $\epsilon_{i}$ a residual error term. 
We then simplified these maximal models. To do this, we calculated the AIC
of the chosen maximal model, as well as the AIC's of a suite of candidate simplified models identified
using R~\citep{R} function dredge from packages MuMIn~\citep{MuMIn}. Simplified models were obtained by
systematically removing all possible combinations of terms from the full model.
The best-fitting model was then determined to be the model with the fewest terms where $\delta$AIC<2.

@@

Next, we examined the form of the scaling relationship between each 
property ($Z$, $G$, and $V$) and $S$. The scaling relationship between $Z$ and 
$S$ has been shown to be a power law \citep{Riede2010} of the form

\begin{equation}
\label{Power}
Z_{i}~\alpha S_{i}^{\beta}
\end{equation}

@@

\begin{equation}
\label{Loglog}
\log{Z_{i}} ~ \log{\alpha} + \beta\log{S_{i}}
\end{equation}

Although these relationships are very similar, they imply different error distributions~\citep{Xiao2011}.
Specifically,~\ref{Power} implies a normally-distributed, additive error and~\ref{LogLog} a lognormal,
multiplicative error. As we have no \emph{a priori} reason to believe that our dataset has one error distribution
over another, we follow the recommendations in \citet{Xiao2011} and compared the two
model formulations explicitly. The model with the error distribution most resembling that observed in the empirical
data was then used to test for potential effects of latitude.

Although scaling relationships between $S$, $G$, and $V$ have not been explicitly examined (but see scaling 
relationships for the standard devaitions of each property in \citet{Riede2010}), we expect that they will follow
power laws similar to that of the relationship between $S$ and $Z$. This is because the links taken into account in
calculating $G$ and $V$ are subsets of the total links included when calculating $Z$. As with $Z$, we explicitly 
compared the error distributions of models for $G$ and $V$ using both the power-law and logarithmic formulations. 
In each case, we used the best-fitting equation as a template when assessing the effect of latitude on scaling with
species richness.
